Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-03-15 ForegroundSimulations/ms.tex
Trim text throughout
From 3b0f6bb (fgsim):
====================

In ms.tex,

(*) Trim text throughout

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $\boldsymbol{k}_\perp$ and $k_\parallel$ are the transverse (on the sky) and line-of-sight (into the sky) wavenumbers respectively, $f_{21}$ is the rest frame frequency of the 21~cm spin-flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. Following \citet{par14},
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f,
\end{align}
which is related to the cosmic volume probed by the instrument. In this paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$, and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2 (h^{-1}$~Mpc$)^3$.

@@

In this paper, we investigate the spectral properties of the proposed dish design through their effects on the resultant foreground delay power spectrum and the constraints they place on attenuation required to suppress reflections between the antenna-feed assembly and antenna pairs to within tolerable limits.

@@

We simulate wide-field visibilities for 19-element HERA from all-sky antenna power pattern and foreground models using the PRISim\footnote{The Precision Radio Interferometry Simulator (PRISim) is publicly available at \url{https://github.com/nithyanandan/PRISim}} software package. The simulations cover 24~hr of observation in {\it drift} mode consisting of 80 accumulations spanning 1080~s each. The total bandwidth is 100~MHz centered on 150~MHz consisting of 256 channels with 390.625~kHz frequency resolution. Models of the antenna power pattern and foregrounds are described below.

@@

In HERA-19 array layout, there are 30 unique baseline vectors and 8 unique baseline lengths. Fig.~\ref{fig:asm-dps-beam-chromaticity-baselines} shows the delay power spectra of foregrounds on the 8 unique baseline lengths obtained with the aforementioned models for power pattern at arbitrary pointings on the sky. In all these panels, the full-band spectrally weighted foreground delay power spectra obtained with achromatic, {\it Airy} and fully chromatic simulated beam patterns are shown in black, red, and blue respectively. The brightening of foreground power near the horizon limits (vertical dotted lines) due to the {\it pitchfork} effect \citep{thy15a,thy15b} is prominently seen in all cases. A clear broadening of spillover-wings outside the horizon limits is seen with increasing chromaticity as the beam is changed from the achromatic to the {\it Airy} to the chromatic simulated model. For instance, the spillover from foreground delay power spectrum obtained with chromatic beam pattern is restricted to $|k_\parallel| \lesssim 0.2\,h$~Mpc$^{-1}$, with the {\it Airy} pattern it is restricted to $|k_\parallel| \lesssim 0.15\,h$~Mpc$^{-1}$, while with the achromatic beam it is restricted to $|k_\parallel| \lesssim 0.12\,h$~Mpc$^{-1}$ even on longest baseline lengths. It is also noted that despite the extreme dynamic range of the spectral weighting function employed, the tail of the foreground spillover at $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ with an {\it Airy} beam pattern is many orders of magnitude higher than that using an achromatic beam while that from the simulated chromatic beam is even higher than from an {\it Airy} pattern by a few orders of magnitude. This signals that the foreground spillover shown are truly limited by intrinsic spectral features in the antenna beam patterns.

@@

Here our goal is to estimate the EoR-foreground dynamic range (ratio of EoR signal power to foreground power) in the presence of antenna beam chromaticity, using HERA as an example. As stated earlier, we estimate this ratio for HERA in the most conservative scenario where we consider only {\it foreground avoidance} under simple spectral weighting scheme. We emphasize that other strategies involving {\it foreground removal} and inverse covariance-based weighting schemes will produce more optimal results than presented here.

In order to avoid signal evolution across the entire bandwidth, we use subbands in which EoR signal evolution is not expected to be significant. These subbands with $B_\textrm{eff}=10$~MHz are defined by spectral weights obtained by convolution of {\it Blackman-Harris} window with itself. These subbands are centered at 150~MHz and 170~MHz to estimate EoR-foreground dynamic range at $z \approx 8.47$ and $z \approx 7.36$ respectively. We use both EoR models 1 and 2 to reduce the dependence of this analysis on any single EoR model.

@@

We define baseline-dependent $k_\parallel$-modes of interest for {\it foreground avoidance} as $|k_\parallel| > k_\parallel^\textrm{FA}(\boldsymbol{b})$, with:
\begin{align}
  k_\parallel^\textrm{FA}(\boldsymbol{b}) &= \frac{2\pi\,f_{21}H_0\,E(z)}{c(1+z)^2}\,\left(\frac{|\boldsymbol{b}|}{c} + \frac{\zeta}{B_\textrm{eff}}\right), \label{eqn:kprll-limit}
\end{align}
where, the first term inside the parenthesis denotes the horizon limit, the second relates to the resolution due to subband bandwidth, $B_\textrm{eff}$, and $\zeta$ denotes a buffer to ensure the $k_\parallel^\textrm{FA}(|\boldsymbol{b}|)$ threshold ``safely'' avoids the main lobe of foreground power.

@@

We reiterate that these estimates are obtained in a conservative scenario wherein only simple spectral weighting technique in the {\it foreground avoidance} limit is used, without any foreground removal or sophisticated weighting strategies using inverse-covariance. Further, these estimates represent worst-case limits demonstrating that HERA-19 and its successors will be extremely sensitive for EoR experiments and will benefit even further with more optimal foreground suppression strategies.

>>>>

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $\boldsymbol{k}_\perp$ and $k_\parallel$ are the transverse (on the sky) and line-of-sight (into the sky) wavenumbers respectively, $f_{21}$ is the rest-frame frequency of the 21~cm spin-flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. Following \citet{par14},
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f,
\end{align}
which is related to the cosmic volume probed by the instrument. In this paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$, and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2 (h^{-1}$~Mpc$)^3$.

@@

In this paper, we investigate the spectral properties of the proposed dish design through their effects on the resultant foreground delay power spectrum and the constraints they place on attenuation required to suppress reflections in the instrument.

@@

We simulate wide-field visibilities for 19-element HERA from all-sky antenna power pattern and foreground models using the PRISim\footnote{The Precision Radio Interferometry Simulator (PRISim) is publicly available at \url{https://github.com/nithyanandan/PRISim}} software package. The simulations cover 24~hr of observation in {\it drift} mode consisting of 80 accumulations spanning 1080~s each. The total bandwidth is 100~MHz centered on 150~MHz consisting of 256 channels with 390.625~kHz frequency resolution. The models used are described below.

@@

In HERA-19 array layout, there are 30 unique baseline vectors and 8 unique baseline lengths. Fig.~\ref{fig:asm-dps-beam-chromaticity-baselines} shows the delay power spectra of foregrounds on the 8 unique baseline lengths obtained with the aforementioned models for power pattern at arbitrary pointings on the sky. In all these panels, the full-band foreground delay power spectra obtained with achromatic, {\it Airy} and fully chromatic simulated beam patterns are shown in black, red, and blue respectively. The brightening of foreground power near the horizon limits (vertical dotted lines) due to the {\it pitchfork} effect \citep{thy15a,thy15b} is prominently seen in all cases. A clear broadening of spillover-wings outside the horizon limits is seen with increasing chromaticity as the beam is changed from the achromatic to the {\it Airy} to the chromatic simulated model. For instance, the spillover from foreground delay power spectrum obtained with chromatic beam pattern is restricted to $|k_\parallel| \lesssim 0.2\,h$~Mpc$^{-1}$, with the {\it Airy} pattern it is restricted to $|k_\parallel| \lesssim 0.15\,h$~Mpc$^{-1}$, while with the achromatic beam it is restricted to $|k_\parallel| \lesssim 0.12\,h$~Mpc$^{-1}$ even on longest baseline lengths. It is also noted that despite the extreme dynamic range of the spectral weighting function employed, the tail of the foreground spillover at $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ with an {\it Airy} beam pattern is many orders of magnitude higher than that using an achromatic beam while that from the simulated chromatic beam is even higher than from an {\it Airy} pattern by a few orders of magnitude. This signals that the foreground spillover shown are truly limited by intrinsic spectral features in the antenna beam patterns.

@@

Here our goal is to estimate the EoR-foreground dynamic range (ratio of EoR signal power to foreground power) in the presence of antenna beam chromaticity, using HERA as an example. % As stated earlier, we estimate this ratio for HERA in the most conservative scenario where we consider only {\it foreground avoidance} under simple spectral weighting scheme. We emphasize that other strategies involving {\it foreground removal} and inverse covariance-based weighting schemes will produce more optimal results than presented here. 
In order to avoid signal evolution across the entire bandwidth, we use subbands in which EoR signal evolution is not expected to be significant. These subbands with $B_\textrm{eff}=10$~MHz are defined by spectral weights obtained by convolution of {\it Blackman-Harris} window with itself. These subbands are centered at 150~MHz and 170~MHz to estimate EoR-foreground dynamic range at $z \approx 8.47$ and $z \approx 7.36$ respectively. We use both EoR models 1 and 2 to reduce the dependence of this analysis on any single EoR model.

@@

We define baseline-dependent $k_\parallel$-modes of interest for {\it foreground avoidance} as $|k_\parallel| > k_\parallel^\textrm{FA}(\boldsymbol{b})$, with:
\begin{align}
  k_\parallel^\textrm{FA}(\boldsymbol{b}) &= \frac{2\pi\,f_{21}H_0\,E(z)}{c(1+z)^2}\,\left(\frac{|\boldsymbol{b}|}{c} + \frac{\zeta}{B_\textrm{eff}}\right), \label{eqn:kprll-limit}
\end{align}
where, the first term inside the parenthesis denotes the horizon limit, the second relates to the resolution due to subband bandwidth, $B_\textrm{eff}$, and $\zeta$ denotes a buffer to ensure the $k_\parallel^\textrm{FA}(|\boldsymbol{b}|)$ threshold ``safely'' avoids the main lobe of foreground power. In this study, we use $\zeta=3.5$.

@@

% We reiterate that these estimates are obtained in a conservative scenario wherein only simple spectral weighting technique in the {\it foreground avoidance} limit is used, without any foreground removal or sophisticated weighting strategies using inverse-covariance. Further, these estimates represent worst-case limits demonstrating that HERA-19 and its successors will be extremely sensitive for EoR experiments and will benefit even further with more optimal foreground suppression strategies.
